Generating urban morphologies at large scales

### 2019-03-15 paper.tex

The quantification of urban form is in itself covered by a vast literature. Recent work have proposed to apply deep learning techniques directly on vector data, such as \cite{2017arXiv170902939M} does for a worldwide classification of road networks. Such an approach avoids the question of isolating relevant features. As we aim at calibrating generative models, we need to have interpretable measures. \cite{boeing2018measuring} proposes an extensive review of existing measures from a large extent of disciplines, their implications for planning and design, and the relation with urban complexity. \citep{webster1995urban} uses image processing techniques such as contrast or Fourier analysis, to extract synthetic descriptions of urban areas from satellite imaging. \cite{fumega2014identification} provides a typology of cities in relation with energy consumption in the perspective of climate change. \cite{rode2014cities} relates indicators of urban form with residential heat-energy demand. Other complexity-related approaches such as fractal dimensions have been introduced as for example by \cite{batty1987fractal}.

@@

The most simple ``realistic'' generator is similar to procedural modeling and distributes building blocks into the space. Given a number $N_B$ of blocks, random positions are drawn and block of a random height and width (with minimal value $m_B$ and maximal value $M_B$ as parameters) are placed at these.

@@

Kernel mixture are a classical way to represent the spatial distribution of population density in an urban area \citep{anas1998urban}. They remain relevant at our scale, as they can be interpreted as a superposition of ``density hotspots'', as can the planning or the self-organization of a district can be. Given a number of centers $N_K$, $\vec{x}_{1\leq j \leq N_K}$ random position are drawn in the grid, at which kernels are applied, such that $s_i = \mathbbm{1}_{d_i \geq \theta_K}$ where the density $d_i$ for the point at position $\vec{x}_i$ is given by
\begin{equation}
    d_i = \frac{1}{N_K}\cdot \sum_j \exp{\left(-\norm{\vec{x}_j - \vec{x}_i}/d_K\right)} 
\end{equation}
where $d_K$ is a range parameter giving the extent of kernels.

@@

The last generator we used is based on network percolation, in the spirit of capturing the constraints imposed by flows traversing a given urban area. While the two previous generator were based on building processes, this one relies on streets, and thus on processes linked to transportation. The idea is to link a fixed number $N_P$ of border points, which can be understood as entrances/exits of the area. Starting with a grid network without links and nodes at a regular spatial sampling (fixed with a step of 5 units in our case), an iterative procedure (i) draws a random number and add a random link at an empty potential link if it is smaller than a parameter called the percolation probability $p_P$; (ii) computes the largest connected component of the network and the number of nodes of this component on the boundary of the world; (iii) stops if this number is equal to the parameter $N_P$. Cells not covered by the resulting giant component are then occupied, at the exception of cells within a neighborhood $L_P$ of a link of the giant component. This way, this component can be understood as a circulating area linking $N_P$ entrances and exits, with a constraint on width through $L_P$.

Note that our generators will be ``fairly compared'' in terms of calibration, as they have the same number of parameters (although we do not introduce any information criteria that would yield the same penalization for overfitting). We show in Fig.~\ref{fig:generators} visual representations of some outputs of each generator.

@@

We compute the morphological indicators given above on a large sample on real urban areas. For practical computational reasons, we restrain our geographical area of study to European urban areas as provided by \cite{}. We expect to already have a good representativity, although not universal, of existing urban forms with this sample, as it is known that European cities already have a significant morphological diversity \citep{le2015forme}. We collect building layouts from OpenStreetMap, as this source has been shown to have a good quality especially in Europe \citep{mooney2010towards}. Using the \texttt{osmosis} tool, buildings are filtered from the openstreetmap raw dump for Europe (downloaded from \url{http://download.geofabrik.de/} in March 2019) and inserted into a Postgis database, which can then be efficiently queried for a specific bounding box. Indeed, although the library developed provides a direct access to the OpenStreetMap API, query limitations do not allow such a systematic sampling.
%Our tool also provide a direct access to the classic and overpass API % not necessary ?
We sample $N=72,000$ points into polygons corresponding to urban areas, first by selecting the area with a uniform selection weighted by population of areas, then by drawing uniform spatial coordinates within the polygon with a polygon sampling heuristic. % TODO clarify the polygon sampler

@@

>>>>

The quantification of urban form is in itself covered by a vast literature.
Recent work have proposed to apply deep learning techniques directly on vector data, such as \cite{2017arXiv170902939M} does, for a worldwide classification of road networks. Such an approach avoids the question of isolating relevant features.
However, as we aim at calibrating generative models, we need to have interpretable measures.
%devrait on justifier pourquoi on a besoin de mesures interprétables ???
\cite{boeing2018measuring} proposes an extensive review of existing measures from a large extent of disciplines, their implications for planning and design, and the relation with urban complexity. \citep{webster1995urban} uses image processing techniques such as contrast or Fourier analysis, to extract synthetic descriptions of urban areas from satellite imaging. \cite{fumega2014identification} provides a typology of cities in relation with energy consumption in the perspective of climate change. \cite{rode2014cities} relates indicators of urban form with residential heat-energy demand. Other complexity-related approaches such as fractal dimensions have been introduced as for example by \cite{batty1987fractal}.

@@

% bon, pour moi les deux premiers modèles sont des modèles procéduraux et pourraient êtres formalisés comme des mpp mais je ne sais pas si c'est pertinent de redire ça à chaque fois...
The most simple ``realistic'' generator is similar to procedural modeling or marked point processes and distributes building blocks into the space (see Fig.~\ref{fig:blocks}). Given a number $N_B$ of blocks, random positions are drawn and block of a random height and width (with minimal value $m_B$ and maximal value $M_B$ as parameters) are placed at these.

@@

Kernel mixture are a classical way to represent the spatial distribution of population density in an urban area \citep{anas1998urban} (see Fig.~\ref{fig:mixture}). They remain relevant at our scale, as they can be interpreted as a superposition of ``density hotspots'', as can the planning or the self-organization of a district can be. Given a number of centers $N_K$, $\vec{x}_{1\leq j \leq N_K}$ random position are drawn in the grid, at which kernels are applied, such that $s_i = \mathbbm{1}_{d_i \geq \theta_K}$ where the density $d_i$ for the point at position $\vec{x}_i$ is given by
\begin{equation}
    d_i = \frac{1}{N_K}\cdot \sum_j \exp{\left(-\norm{\vec{x}_j - \vec{x}_i}/d_K\right)} 
\end{equation}
where $d_K$ is a range parameter giving the extent of kernels.
% je ne suis pas convaincu par la formule : il y a un truc que j'ai raté ou il manque un vecteur
% si les x_i correspondent aux posistion des centres (hotspots), alors on ne calcule les densités que sur les hotspots et il nous manque la possibilité de calculer les densités pour les autres points. On aurait pas besoin de 2 vecteurs ? x_i => les centres, y_i => l'ensemble des points ?
% Dans le papier d'Anas il triche un peu : il utilise x_ij (mn dans le papier) pour la distance entre le point i et le centre j...

@@

The last generator we used is based on network percolation, in the spirit of capturing the constraints imposed by flows traversing a given urban area (see Fig.~\ref{fig:percolation}). While the two previous generator were based on building processes, this one relies on streets, and thus on processes linked to transportation. The idea is to link a fixed number $N_P$ of border points, which can be understood as entrances/exits of the area. Starting with a grid network without links and nodes at a regular spatial sampling (fixed with a step of 5 units in our case), an iterative procedure (i) draws a random number and add a random link at an empty potential link if it is smaller than a parameter called the percolation probability $p_P$; (ii) computes the largest connected component of the network and the number of nodes of this component on the boundary of the world; (iii) stops if this number is equal to the parameter $N_P$. Cells not covered by the resulting giant component are then occupied, at the exception of cells within a neighborhood $L_P$ of a link of the giant component. This way, this component can be understood as a circulating area linking $N_P$ entrances and exits, with a constraint on width through $L_P$.

Note that our generators will be ``fairly compared'' in terms of calibration, as they have the same number of parameters (although we do not introduce any information criteria that would yield the same penalization for overfitting).
We show in Fig.~\ref{fig:generators} visual representations of some outputs of each generator, including a random generator (see Fig.~\ref{fig:random}).

@@

We compute the morphological indicators given above on a large sample on real urban areas. For practical computational reasons, we restrain our geographical area of study to European urban areas as provided by \cite{}. We expect to already have a good representativity, although not universal, of existing urban forms with this sample, as it is known that European cities already have a significant morphological diversity \citep{le2015forme}. We collect building layouts from OpenStreetMap (illustrated in Figure~\ref{fig:osm}), as this source has been shown to have a good quality especially in Europe \citep{mooney2010towards}. Using the \texttt{osmosis} tool, buildings are filtered from the openstreetmap raw dump for Europe (downloaded from \url{http://download.geofabrik.de/} in March 2019) and inserted into a Postgis database, which can then be efficiently queried for a specific bounding box. Indeed, although the library developed provides a direct access to the OpenStreetMap API, query limitations do not allow such a systematic sampling.
%Our tool also provide a direct access to the classic and overpass API % not necessary ?
We sample $N=72,000$ points into polygons corresponding to urban areas, first by selecting the area with a uniform selection weighted by population of areas, then by drawing uniform spatial coordinates within the polygon with a polygon sampling heuristic. % TODO clarify the polygon sampler

@@

%%%%%%%%%%%%%%
\begin{figure}
    \centering
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(2.767548582324828,42.12774259047379)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(2.956412030079216,48.95833438674378)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(2.8534915107882908,48.87943398394085)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(4.724576799597424,46.11247464316661)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(6.693161940127149,51.28067900621835)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(7.665417780766332,51.61719462734998)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(7.746400123854242,49.448716515580514)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(8.66720808445827,49.469125872820456)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(8.368921085687871,49.224787816959754)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(9.3988783289936,45.52643666558245)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(9.422039749719488,48.68506062982569)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(10.529303821177011,57.458935658684915)}.png}
    \end{subfigure}
    \caption{Example of patterns extracted from OpenStreetMap.}
    % TODO should provide parameters here
    \label{fig:osm}
\end{figure}
%%%%%%%%%%%%%%
